---
output:
  word_document: default
  html_document: default
---

### Random Forests

```{r}
options(tidyverse.quiet = TRUE)
library(tidyverse)
library(caret)
library(ranger)
```


```{r}
Blood <- read_csv("C:/Users/timbo/Documents/BAN502-RmdData/Blood.csv")
```

```{r}
str(Blood)
summary(Blood)
```

Convert the DonatedMarch variable to a factor and recode the variable  
```{r}
Blood <- Blood %>% mutate(DonatedMarch = as_factor(as.character(DonatedMarch))) %>%
  mutate(DonatedMarch = fct_recode(DonatedMarch,
                             "No" = "0",
                             "Yes" = "1"))
str(Blood)
```

Task 1 - Split the dataset into training (70%) and testing (30%) sets
```{r}
set.seed(1234) 
train.rows = createDataPartition(y = Blood$DonatedMarch, p=0.7, list = FALSE) 

train <- slice(Blood, train.rows)
test <- slice(Blood, -train.rows)
```

Task 2 - Create a random forest model on the training set to predict DonatedMarch
```{r}
fit_control <- trainControl(method = "cv",  
                           number = 10) 

set.seed(123)  
rf_fit <- train(x=as.data.frame(train[,-5]), y=train$DonatedMarch, 
                  method = "ranger", 
                  importance = "permutation", 
                  trControl = fit_control,
                  num.trees = 100)
```

Task 3 - Use varImp  
What is the most important variable in the model?
TotalDonations is the most important variable in the model.
What is the least important variable in the model?
Mnths_Since_Last is the least important variable in the model.
```{r}
varImp(rf_fit)
rf_fit
```
Task 4 - Use the model to develop predictions on the training set
```{r}
RFpred_train <- predict.train(rf_fit, train)
head(RFpred_train)
```
Task 5 - Use the model to create a confusion matrix using caret's confusionMatrix
What is the accuracy, sensitivity, and specificity of the model?
The accuracy of the model equals 90%, which is better than the naive accuracy of 76%.
The sensitivity of the model equals 61% and the specificity equals 99%.
```{r}
confusionMatrix(RFpred_train, train$DonatedMarch, positive = "Yes")
```

Task 6  
How does the accuracy of the model compare to the naive model that assumes that all observations are in the majority class?
The naive accuracy of the model at 76% is much lower than the accuracy, which is 90%, so the model is a much better model than the naive accuracy.

Task 7 - Use the model to develop predictions on the testing set
```{r}
RFpred_test <- predict.train(rf_fit, test)
head(RFpred_test)
```
Task 7 - Develop a confusion matrix
How does the model perform on the testing set?
The accuracy of the model equals 74%, the sensitivity equals 24% and the specificity equals 90%.
The naive accuracy at 76% is better than the models accuracy, so this is not a very good model. The model on the testing set does not perform well.
```{r}
confusionMatrix(RFpred_test, test$DonatedMarch, positive = "Yes")
```

Task 8  
Comment on how this model might be used in the "real-world."
This model might be used for building finance and investment data models.
Would you recommend this model for real-world use?
Yes, I would recommend this model for "real-world" use.
