---
output:
  word_document: default
  html_document: default
---

### Missing Data  

## Task 1  
Are there students that are missing multiple pieces of data?
Yes, Tutorial is missing 1% of its data, Midterm is missing 3% of its data, TakeHome is missing 3% of its data, and Final is missing 4% of its data.  

## Task 2  
How much data is missing and in what variables?
There is one case missing Tutorial, one case missing both Midterm and TakeHome, two cases missing Midterm alone, two cases missing only TakeHome, and four cases missing only Final.  

## Task 3  
Use row-wise deletion of missing values. How many rows remain in this data frame?
After using row-wise there are now 89 rows remaining.  

## Task 4  
Use column-wise deletion of missing values from the original data frame. How many columns remain in this data frame?
After using column-wise there are now 2 columns remaining.  

## Task 5  
Which approach (Task 3 or Task 4) seems preferable for this dataset? 
I believe the row-wise approach is the most preferable for this dataset because after performing row-wise there are six columns left to build your model(s) with as opposed to only having two columns after performing the column-wise approach.  

## Task 7  
Briefly discuss potential issues that could be encountered when working with missing data. Describe situations where imputation may not be advisable.
Missing data can negatively influence a model and create biased estimates leading to invalid outcomes and present the following problems:
* reduce the "probability the test will reject the null hypothesis when it is false"
* cause bias in the "estimation of parameters"
* reduce the "representation of the samples", and
* could "complicate the analysis of the study"
(source https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3668100/)

```{r echo=FALSE}
options(tidyverse.quiet=TRUE)
library(tidyverse)
library(VIM)
library(mice)
```

```{r}
class_grades <- read_csv("C:/Users/timbo/Documents/BAN502-RmdData/class-grades.csv")

```

```{r}
grades <- read_csv("C:/Users/timbo/Documents/BAN502-RmdData/class-grades.csv")
str(grades)
summary(grades)
```


```{r}
vim_plot <- aggr(grades, numbers = TRUE, prop = c(TRUE, FALSE), cex.axis = .7)
```

```{r}
grades <- grades %>% drop_na(Tutorial)
grades <- grades %>% drop_na(Midterm)
grades <- grades %>% drop_na(TakeHome)
grades <- grades %>% drop_na(Final)
```

```{r}
vim_plot <- aggr(grades, numbers = TRUE, prop = c(TRUE, FALSE), cex.axis = .7)
```

```{r}
grades <- class_grades %>% select(-Tutorial)
grades <- class_grades %>% select(-Midterm)
grades <- class_grades %>% select(-TakeHome)
grades <- class_grades %>% select(-Final)
```

```{r}
vim_plot <- aggr(grades, numbers = TRUE, prop = c(TRUE, FALSE), cex.axis = .7)
```
## Task 6  
```{r}
grades_imp <- mice(grades, m=1, method = 'pmm', seed = 12345)
summary(grades_imp)

```
## Task 6  
```{r}
densityplot(grades_imp)
```
## Task 6  
```{r}
grades_complete <- complete(grades_imp)
summary(grades_complete)

```

