---
output:
  word_document: default
  html_document: default
---

### Classification Trees  

```{r echo=FALSE}
library(tidyverse)
library(caret)
library(rpart)
library(rattle)
library(RColorBrewer)
```

```{r}
parole <- read_csv("C:/Users/timbo/Documents/BAN502-RmdData/parole.csv")
```
```{r}
parole <- parole %>% mutate(male = as_factor(as.character(male))) %>%
  mutate(male = fct_recode(male,
                             "Female" = "0",
                             "Male" = "1"))

parole <- parole %>% mutate(race = as_factor(as.character(race))) %>%
  mutate(race = fct_recode(race,
                             "White" = "1",
                             "Otherwise" = "2"))

parole <- parole %>% mutate(state = as_factor(as.character(state))) %>%
  mutate(state = fct_recode(state,
                             "Any other state" = "1",
                             "Kentucky" = "2",
                             "Louisiana" = "3",
                             "Virginia" = "4"))

parole <- parole %>% mutate(crime = as_factor(as.character(crime))) %>%
  mutate(crime = fct_recode(crime,
                             "Any other crime" = "1",
                             "Larceny" = "2",
                             "Drug-related crime" = "3",
                             "Driving-related crime" = "4"))

parole <- parole %>% mutate(multiple.offenses = as_factor(as.character(multiple.offenses))) %>%
  mutate(multiple.offenses = fct_recode(multiple.offenses,
                             "Otherwise" = "0",
                             "Incarcerated for multiple offenses" = "1"))

parole <- parole %>% mutate(violator = as_factor(as.character(violator))) %>%
  mutate(violator = fct_recode(violator,
                             "Completed parole without violation" = "0",
                             "Violated parole" = "1"))
```

Task 1 - Split the parole data into training and testing sets  
```{r}
train.rows <- createDataPartition(y = parole$violator, p=0.7, list = FALSE)
set.seed(12345)

train <- slice(parole, train.rows)
test <- slice(parole, -train.rows)
```

Task 2 - Create a classification tree  
```{r}
tree1 <- rpart(violator~., train, method="class")
fancyRpartPlot(tree1, tweak = .4, cex=.8)
```

Task 3  
For the tree created in Task 2, how would you classify a 40-year-old parolee from Louisiana who served a 5 year prison sentence? Describe how you walk through the classification tree to arrive at the answer.
Starting at the top of the tree at "state = Any other state, Kentucky, Virginia," and knowing that our parolee is from Louisiana, you should choose "no" which leads you to "multiple.offenses = otherwise." Our parolee did not have multiple offenses, so you would move to the left of the tree, which then leads you to "time.served < 4.8." Our parole's time served equals 5 years, so we would choose "no," and you would move right, which takes you to "Violated parole." Our 40-year-old parolee from Louisiana, who served a 5 year sentence did violate his parole.

Task 4 - Use the printcp function to evaluate tree performance  
What cp value should be selected?
The cp value that should be selected is 0.01.
```{r}
printcp(tree1)
plotcp(tree1)
```

Task 5  
Which class (category) in the training set is the majority class (ie., has the most observations)?
The class in the training set that has the most observations is "Completed parole without violation.
```{r}
tree2 <- prune(tree1, cp = tree1$cptable[which.min(tree1$cptable["xerror"]), "CP"])

fancyRpartPlot(tree2, tweak = .4, cex=.8)
```

Task 6 - Develop predictions for the training data  
```{r}
treepred1 <- predict(tree1, train, type = "class")
head(treepred1)
```
Task 6 - Use caret's confustionMatrix function to calculate the accuracy, specificity, and sensitivity  
```{r}
confusionMatrix(treepred1,train$violator,positive="Completed parole without violation")
```
Task 7 - Develop predictions for the testing data  
```{r}
treepred2 <- predict(tree1, test, type = "class")
head(treepred2)
```
Task 7 - Use caret's confustionMatrix function to calculate the accuracy, specificity, and sensitivity  
```{r}
confusionMatrix(treepred2,test$violator,positive="Completed parole without violation")
```
Task 8 - Read in the Blood dataset   
```{r}
Blood <- read_csv("C:/Users/timbo/Documents/BAN502-RmdData/Blood.csv")
```
Task 8 - Convert the DonatedMarch variable to a factor  
```{r}
Blood <- Blood%>% mutate(DonatedMarch = as_factor(as.character(DonatedMarch))) %>%
  mutate(DonatedMarch = fct_recode(DonatedMarch,
                             "No" = "0",
                             "Yes" = "1"))
```

Task 9 - Split the Blood dataset into training (70%) and testing (30%) sets  
```{r}
test.rows <- createDataPartition(y = Blood$DonatedMarch, p=0.7, list = FALSE)
set.seed(1234)

train2 <- slice(Blood, test.rows)
test2 <- slice(Blood, -test.rows)
```


Task 9 - Develop a classification tree to predict "DonatedMarch" on the training set
```{r}
train_tree1 <- rpart(DonatedMarch~., train2, method="class")
fancyRpartPlot(train_tree1, tweak = .9, cex=.6)
```

Task 9 - Evaluate the complexity parameter (cp) selection for this model  
```{r}
printcp(train_tree1)
plotcp(train_tree1)
```

Task 10 - Prune the tree back to the optimal cp value on the training set
```{r}
train_tree2 <- prune(train_tree1, cp = train_tree1$cptable[which.min(train_tree1$cptable["xerror"]), "CP"])

fancyRpartPlot(train_tree2, tweak = .9, cex=.6)
```

Task 10 - Make predictions on the training set
```{r}
treepred2 <- predict(train_tree2, train2, type = "class")
head(treepred2)
```

Task 10 - Use the confustionMatrix function on the training set
```{r}
confusionMatrix(treepred2,train2$DonatedMarch,positive="Yes")
```

Task 10 - Prune the tree back to the optimal cp value on the testing set
```{r}
test_tree1 <- rpart(DonatedMarch~., test2, method="class")
fancyRpartPlot(test_tree1)
```

Task 10 - Make predictions on the testing set
```{r}
treepred3 <- predict(test_tree1, test2, type = "class")
head(treepred3)
```

Task 10 - Use the confustionMatrix function on the testing set
```{r}
confusionMatrix(treepred3,test2$DonatedMarch,positive="Yes")
```

