---
output:
  word_document: default
  html_document: default
---
### Module 2 - Assignment 2

## Langdon, Christy

# Task 1
Comment as to why we convert "yr", "mnth", and "hr" into factors?
It is a good idea to convert "yr", "mnth", and "hr" into factors to rename the levels of the factors in order to be more descriptive.

Why not just leave them as numbers?
Its descriptive nature creates better communique with the end user when explaining results.

**You many find it useful to complete some of the other tasks and then come back to answer this question.


# Task 2
Which of the quantative variables appears to be best correlated with "count"?
**ignore the "registered" and "casual" variable as the sum of these two variables equals "count."

Atemp (Normalized feeling temperature in Celsius) appears to be the best correlated with "count."

# Task 3
Conduct a boxplot-based analysis for each of the categorical variables.
Which variables appear to affect "count"?



Provide a brief explanation as to why you believe that each variable does or does not affect "count" (use your intution to help you answer this question).
**Use gridarrange to reduce the space needed for plots

# Task 4
Use forward stepwise regression to build a multiple linear regression model to predict "count."

What variables are included in your forward model?
The following variables are included in the forward model; season, yr, mnth, hr, holiday, weekday, workingday, weathersit, temp, atemp, hum, windspeed, and count.

Comment on the quality of the model.
Does the model match our intuition/common sense?
Is there evidence of multicollinearity?


# Task 5
Does this model differ from the forward model?


# Task 6
Describe how "workingday" is represented in the model via other variables.


# Task 7
Comment on the usability of this model. Any cautions concerning its potential use?

```{r}
options(tidyverse.quiet = TRUE)
library(tidyverse)
library("GGally")
library(MASS)
library(car)
library(ggplot2)
library(gridExtra)

```

```{r}
library(readr)
bike <- hour <- read_csv("C:/Users/timbo/Documents/BAN502-RmdData/hour.csv")
View(hour)

```
```{r}
bike <- bike %>% mutate(season = as_factor(as.character(season))) %>%
  mutate(season = fct_recode(season,
                             "Spring" = "1",
                             "Summer" = "2",
                             "Fall" = "3",
                             "Winter" = "4"))

```

```{r}
bike <- bike %>% mutate(yr = as_factor(yr))
str(bike)

```

```{r}
bike <- bike %>% mutate(mnth = as_factor(mnth))
str(bike)

```

```{r}
bike <- bike %>% mutate(hr = as_factor(hr))
str(bike)

```

```{r}
bike <- bike %>% mutate(holiday = as_factor(as.character(holiday))) %>%
  mutate(holiday = fct_recode(holiday,
                             "NotHoliday" = "0",
                             "Holiday" = "1"))

```

```{r}
bike <- bike %>% mutate(workingday = as_factor(as.character(workingday))) %>%
  mutate(workingday = fct_recode(workingday,
                             "NotWorkingDay" = "0",
                             "WorkingDay" = "1"))
```

```{r}
bike <- bike %>% mutate(weathersit = as_factor(as.character(weathersit))) %>%
  mutate(weathersit = fct_recode(weathersit,
                             "NoPrecip" = "1",
                             "Misty" = "2",
                             "LightPrecip" = "3",
                             "HeavyPrecip" = "4"))

```

```{r}
bike <- bike %>% mutate(weekday = as_factor(as.character(weekday))) %>%
  mutate(weekday = fct_recode(weekday,
                             "Sunday" = "0",
                             "Monday" = "1",
                             "Tuesday" = "2",
                             "Wednesday" = "3",
                             "Thursday" = "4",
                             "Friday" = "5",
                             "Saturday" = "6"))

```


```{r}
season <- ggplot(bike, aes(season, count)) + geom_boxplot()

holiday <- ggplot(bike, aes(holiday, count)) + geom_boxplot()

weekday <- ggplot(bike, aes(weekday, count)) + geom_boxplot()

workingday <- ggplot(bike, aes(workingday, count)) + geom_boxplot()

weathersit <- ggplot(bike, aes(weathersit, count)) + geom_boxplot()

library(gridExtra)
grid.arrange(season, holiday, weekday, workingday, weathersit, ncol = 2)

```

```{r}
bike2 <- bike %>%
  dplyr::select(-c("instant", "dteday", "registered", "casual"))

```

```{r}
summary(bike2)
glimpse(bike2)

```

```{r}
ggplot(bike2, aes(x=count)) +
  geom_histogram() +
  theme_bw()

```

```{r}
ggpairs(bike2, columns = c("season", "holiday", "weekday", "count"))
ggpairs(bike2, columns = c("workingday", "weathersit", "temp", "windspeed", "count"))

```

```{r The first model uses the variable that is best correlated with "count"}
ggplot(data = bike) +
  geom_point(mapping = aes(y = count, x = windspeed)) +
  theme_bw()

```

```{r}
model1 <- lm(count ~ windspeed, bike2)
summary(model1)
ggplot(bike2, aes(x=windspeed, y=count)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  theme_bw()

```

```{r Create all in model}
allmodel <- lm(count ~., bike2)
summary(allmodel)

```

```{r Create a model with all possible predictors}
emptymodel <- lm(count ~1, bike2)
summary(emptymodel)

```

```{rd Backward stepwise}
backmodel <- stepAIC(allmodel, direction = "backward", trace = TRUE)
summary(backmodel)

```

```{r}
forwardmodel <- stepAIC(emptymodel, direction = "forward", scope = list(upper=allmodel, lower=emptymodel,
                        trace = TRUE))
summary(forwardmodel)
                        
```

```{r}
allmodel <- lm(count ~., bike2)
summary(allmodel)
```

```{r}
ggplot(bike2, aes(x=weathersit, y=count)) + geom_point() +
  theme_bw()
```

