
```{r}
options(tidyverse.quiet = TRUE)
library(tidyverse)
library(cluster)
library(factoextra)
library(dendextend)
```

```{r}
trucks <- read_csv("C:/Users/timbo/Documents/BAN502-RmdData/trucks.csv")
```

```{r}
trucks_cluster <- scale(trucks)
summary(trucks_cluster)
```
Task 1 - Plot the relationship between Distance and Speeding
There appears to be a natural clustering of Distance equaling between 25 and 75, then also 150 to 200.
```{r}
ggplot(trucks, aes(x=Distance, y=Speeding)) +
  geom_point() +
  theme_bw()
```
Task 2 - Create a new data frame that excludes Driver_ID and includes scaled versions  of the Distance and Speeding variables
```{r}
trucks2 <- trucks %>%
  dplyr::select(-c("Driver_ID"))

trucks2 <- as.data.frame(scale(trucks2)) 
summary(trucks2)
```
Task 3- Use k-Means with two clusters to cluster the trucks2 data frame

There are two clusters. 
```{r}
set.seed(64)
cluster1 <- kmeans(trucks2, 2)
```

```{r}
fviz_cluster(cluster1, trucks2)
```
Task 4 - Use the two methods from k-Means lecture to identify the optimal number of clusters
```{r}
set.seed(64)
fviz_nbclust(trucks2, kmeans, method = "wss")
```

```{r}
set.seed(64)
fviz_nbclust(trucks2, kmeans, method = "silhouette")
```
Task 5 - Use the optimal number of clusters to create k-Means clusters
```{r}
set.seed(64)
cluster2 <- kmeans(trucks2, 4)
cluster2 
fviz_cluster(cluster2, trucks2)
```

Task 6  
In words, how would you characterize the clusters you created in Task 5?

Read in the kenpom20 file into a data frame called bball
```{r}
bball <- read_csv("C:/Users/timbo/Documents/BAN502-RmdData/kenpom20.csv")
```

Task 7- Create a new data frame called bball2 that excludes TeamName and scales the variables
Is there consensus between between these two methods as the optimal number of clusters?
```{r}
bball2 <- bball %>%
  dplyr::select(-c("TeamName"))

bball2 <- as.data.frame(scale(bball2)) 
summary(bball2)
```

```{r}
set.seed(123)
cluster3 <- kmeans(bball2, 4)
```

```{r}
fviz_cluster(cluster3, bball2)
```

```{r}
set.seed(123)
fviz_nbclust(bball2, kmeans, method = "wss")
```

```{r}
set.seed(123)
fviz_nbclust(bball2, kmeans, method = "silhouette")
```
Task 8 - Create k-Means clusters 
```{r}
set.seed(1234)
cluster3 <- kmeans(bball2, 4)
```

```{r}
fviz_cluster(cluster3, bball2)
```

```{r}
cluster <- as.data.frame(cluster3$cluster)
clusternum <- bind_cols(bball2,cluster)
str(bball2)
```



```{r}
bball2 <- bball2 %>% mutate(clusternum = cluster3$cluster)
```


```{r}
ggplot(bball, aes(x=AdjOE, y=AdjDE, color=factor(clusternum))) + 
  geom_point() + 
  facet_wrap(~factor(cluster))
```

